---
title: "Untitled"
output:
  html_document:
    code_folding: "hide"
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## packages

J'ouvre un nouveau chunk avec : Ctrl + Alt + I  
Je charge la liste des packages nécessaires à la formation.

```{r}
# Load libraries ---------------------------------------------------------------
library(sp)
library(rgdal)
library(raster)
library(maps)
library(mapdata)
library(dplyr)
library(here)
# Other libraries ----
library(lattice) #to improve graphics
library(MASS)
library(ggplot2)
# devtools::install_github('bhaskarvk/widgetframe')
library(leaflet)
library(ggmap)

```

## Dossier de travail

```{r}
# Define working directories ---------------------------------------------------
# WD <- "/mnt/Data/Formation_TD-GLM_Cours_Agro/TD_Vilaine_SIG-et-GLM_2017_01_3days"
WD <- here()
# Folder of original files
origWD <- here("01_Original_data")
# Folder for outputs
saveWD <- here("02_Outputs")
# Folder where to save outputs from R
figWD <- here("03_Figures")
# Folder where complementary functions are stored
funcWD <- here("04_Functions")
# Copy maintained complementary functions from local directory
# if (file.exists("~/Rshiny/Map_creation_git/Rsources/Map.output.fun.R")) {
#   tmp <- file.copy("~/Rshiny/Map_creation_git/Rsources/Map.output.fun.R",
#                    paste0(funcWD, "/Map.output.fun.R"), overwrite = TRUE)
# }
# Manual operations
Manuel <- FALSE


```

## Scripts supplémentaires

```{r}
# Download other personnal graphic functions
source(paste0(funcWD, "/function.Gamma.Hist.R"))
source(paste0(funcWD, "/function.chisq.gof.R"), encoding = "Latin1")
```

## Les données

### Lecture
Les données sont des données spatialisées.

```{r}
# Get data --------------------------------------------------------------------
# Load Coastlines shapefile
# Coasts_Vilaine_wgs84 <- readOGR(dsn = saveWD, layer = "Dept_Contour_wgs84") 
Coasts_Vilaine_wgs84 <- readr::read_rds(file.path(saveWD, "Dept_Contour_wgs84.rds"))
# Coasts_Vilaine_wgs84 <- readr::write_rds(Coasts_Vilaine_wgs84, file.path(saveWD, "Dept_Contour_wgs84.rds"))
# Load the data shapefile
# dataset <- readOGR(dsn = saveWD, layer = "Stations_covariates_wgs84")
Stations_covariates_wgs84 <- readr::read_rds(
  path = paste0(saveWD, "/Stations_covariates_wgs84.rds"))

# Plot data
plot(Stations_covariates_wgs84)
plot(Coasts_Vilaine_wgs84, usePolypath = F, lwd = 2, col = "grey", add = TRUE)
box()
```

### Préparation

```{r}
# Transform to non-spatial dataset
dataset.tmp <- as.tbl(Stations_covariates_wgs84@data)
dataset.tmp
dim(dataset.tmp)
summary(dataset.tmp)
names(dataset.tmp)

# Clean dataset
names(dataset.tmp) <- gsub("_pt", "", names(dataset.tmp))

# Add a new covariate: bathymetry as class factor covariate
# c("1depth < 5 m", "2depth 5 - 10 m", "3depth 10 - 20 m", "4depth 20 - 50 m")
dataset <- dataset.tmp %>%
  mutate(Bathy_c = 
           cut(-Bathy, breaks = c(min(-Bathy), 5, 10, 20, max(-Bathy, 50)),
               labels = c("1depth < 5m", "2depth 5-10m",
                          "3depth 10-20m", "4depth 20-50m"),
               include.lowest = TRUE)) %>%
  # dplyr::select(-matches("coords"), -optional, -id) %>%
  dplyr::select(lat:Density, Sedim, Bathy, Bathy_c) %>%
  mutate(Year = as.factor(as.character(Year)))

rm(dataset.tmp)
```

## Exploration des distributions

```{r}
ggplot(dataset) +
  geom_histogram(aes(log(Density + 1)),
                 bins = 30, fill = "orange", col = "grey40") +
  facet_grid(Bathy_c ~ Sedim, scales = "free_y")
```

### Comparaison des distributions 

```{r}
# _Distribution type of the data ----
# Tests of normality
par(mfrow = c(1,2))
qqnorm(dataset$Density) ; qqline(dataset$Density, col = "red")
qqnorm(log(dataset$Density + 1)) ; qqline(log(dataset$Density + 1), col = "red")

ks.test(rnorm(1000, 0, 1), pnorm)
ks.test(dataset$Density, pnorm)
ks.test(log(dataset$Density + 1), pnorm)

chisq.gof(rnorm(1000, 0, 1), distribution = "normal")
chisq.gof(log(dataset$Density + 1), distribution = "normal")
# Comments : these tests are to be used with caution (need lot of data)

# Gamma ?
Gamma.Hist(dataset$Density + 1)

shape <- mean(dataset$Density) * mean(dataset$Density) / var(dataset$Density)
chisq.gof(dataset$Density, distribution = "gamma", shape = shape)

```

### Distribution selon les covariables

```{r}
# _Distribution with covariates ----
# qqplot
x11()
par(mfrow = c(2, length(levels(as.factor(dataset$Bathy_c)))))
tapply(dataset$Density, dataset$Bathy_c,
       function(x) {qqnorm(x); qqline(x, col = "red")})
tapply(log(dataset$Density + 1), dataset$Bathy_c, 
       function(x) {qqnorm(x); qqline(x, col = "red")})

# Statistical tests
tapply(log(dataset$Density + 1), dataset$Bathy_c,
       function(x) {ks.test(x, y = pnorm)})
tapply(log(dataset$Density + 1), dataset$Bathy_c,
       function(x) {chisq.gof(x, distribution = "normal")})

# Histograms with Gamma distribution
par(mfcol = c(1, length(levels(as.factor(dataset$Bathy_c)))))
tapply(dataset$Density + 1, dataset$Bathy_c, Gamma.Hist)

```

### Conclusion intermédiaire

- Le modèle ne sera sûrement pas gaussien, peut-être Gamma ou Log-normal
- On a quand même un problème avec les absences

## Exploration des interactions

```{r}
# log density
par(mfrow = c(2, 2))
interaction.plot(x.factor = dataset$Bathy_c,  trace.factor = dataset$Year,
                 response = log(dataset$Density + 1))
interaction.plot(x.factor = dataset$Year,  trace.factor = dataset$Bathy_c,
                 response = log(dataset$Density + 1))
interaction.plot(x.factor = dataset$Sedim, trace.factor = dataset$Year,
                 response = log(dataset$Density + 1),  col = rainbow(15))
interaction.plot(x.factor = dataset$Sedim, trace.factor = dataset$Bathy_c,
                 response = log(dataset$Density + 1) )

```

## Ajustement et sélection de modèles

 REMEMBER ====------------------------------------------------------------
 Define what you are looking for, which question you try to answer!

 MODEL TESTING
 Test different form of models regarding combination of covariates and distribution form of the residuals;
 Compare models using statistical tools (AIC, anova, …), cross-validation tools but also your knowledge of the dataset and the answers you are looking for;
 Analyse the residuals. Analyse their distribution and if your hypotheses are verified.

 REMEMBER
 Only when hypotheses on distribution law are verified, physical covariates and interactions selected may start to be interpreted...

 ==== WARNING ====
 Scripts below are just giving some examples, but are not playing best solutions !
 Try other solutions, build your own ! 


```{r}
# Try simple Linear models ----
Dens1.lm <- lm(Density ~ Bathy + Sedim, data = dataset)
summary.glm <- summary(Dens1.lm)
summary.glm

# Diagnostic of the residuals of Dens1.lm
par(mfcol = c(2,2))
plot(Dens1.lm, which = 1:3)
hist(resid(Dens1.lm), main = "Hist. of residuals")

# Analysis of variance
aov1 <- anova(Dens1.lm, test = "Chisq") 
aov1

AIC(Dens1.lm)
```


### Sélection sur la base de l'AIC
```{r, results="hide"}
# Using step AIC
null.m <- glm((Density) ~ 1, data = dataset, family = "Gamma")
Dens3.glm <- stepAIC(null.m, 
                     (Density) ~ (Bathy_c + Sedim + Year)^2,
                     data = dataset, direction = "both")
```

Best model
```{r}
model <- glm(log(Density+1) ~ Bathy_c + Sedim, family = gaussian, data = dataset)
AIC(model) + 2 * sum(log(dataset$Density + 1))

plot(model)
```

### Prédictions
```{r}
# Predictions ----
# Dowload the file with factors for predictions
predictions <- readr::read_delim(paste0(origWD, "/predictions.csv"),
                                 delim = ";")

# Compute predictions (in the scale of the response)
pred.Density <- cbind(predictions, predict.glm(model, predictions, type = "r", se = T)) %>%
  group_by(Sedim, Bathy_c) %>%
  summarise(mean.fit = mean(fit),
            mean.se = mean(se.fit)) %>%
  ungroup()

# library(emojifont)
ggplot(pred.Density, aes(x = Bathy_c, y = mean.fit, colour = Sedim)) +
  geom_point() +
  #geom_text(family = 'EmojiOne', size = 8, label = emoji('fish'),
  #          show.legend = FALSE) + 
  geom_errorbar(aes(ymin = mean.fit - mean.se, ymax = mean.fit + mean.se), 
                width = 0.2) +
  theme(legend.position = "bottom")


```

