---
title: "Modèle Delta"
author: "Sébastien Rochette, ThinkR"
output:
  html_document:
    code_folding: "hide"
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## packages

J'ouvre un nouveau chunk avec : Ctrl + Alt + I  
Je charge la liste des packages nécessaires à la formation.

```{r, message=FALSE}
# Load libraries ---------------------------------------------------------------
library(sp)
library(rgdal)
library(raster)
library(maps)
library(mapdata)
library(dplyr)
library(here)
# Other libraries ----
library(lattice) #to improve graphics
library(MASS)
library(ggplot2)
# devtools::install_github('bhaskarvk/widgetframe')
library(leaflet)
library(ggmap)
library(pROC)
```

## Dossier de travail

```{r}
# Define working directories ---------------------------------------------------
# WD <- "/mnt/Data/Formation_TD-GLM_Cours_Agro/TD_Vilaine_SIG-et-GLM_2017_01_3days"
WD <- here()
# Folder of original files
origWD <- here("01_Original_data")
# Folder for outputs
saveWD <- here("02_Outputs")
# Folder where to save outputs from R
figWD <- here("03_Figures")
# Folder where complementary functions are stored
funcWD <- here("04_Functions")
# Copy maintained complementary functions from local directory
# if (file.exists("~/Rshiny/Map_creation_git/Rsources/Map.output.fun.R")) {
#   tmp <- file.copy("~/Rshiny/Map_creation_git/Rsources/Map.output.fun.R",
#                    paste0(funcWD, "/Map.output.fun.R"), overwrite = TRUE)
# }
# Manual operations
Manuel <- FALSE


```

## Scripts supplémentaires

```{r}
# Download other personnal graphic functions
source(paste0(funcWD, "/function.Gamma.Hist.R"))
source(paste0(funcWD, "/function.chisq.gof.R"), encoding = "Latin1")
```

## Les données

### Lecture
Les données sont des données spatialisées.

```{r}
# Get data --------------------------------------------------------------------
# Load Coastlines shapefile
# Coasts_Vilaine_wgs84 <- readOGR(dsn = saveWD, layer = "Dept_Contour_wgs84") 
Coasts_Vilaine_wgs84 <- readr::read_rds(file.path(saveWD, "Dept_Contour_wgs84.rds"))
# Coasts_Vilaine_wgs84 <- readr::write_rds(Coasts_Vilaine_wgs84, file.path(saveWD, "Dept_Contour_wgs84.rds"))
# Load the data shapefile
# dataset <- readOGR(dsn = saveWD, layer = "Stations_covariates_wgs84")
Stations_covariates_wgs84 <- readr::read_rds(
  path = paste0(saveWD, "/Stations_covariates_wgs84.rds"))

# Plot data
plot(Stations_covariates_wgs84)
plot(Coasts_Vilaine_wgs84, usePolypath = F, lwd = 2, col = "grey", add = TRUE)
box()
```

### Préparation

```{r}
# Transform to non-spatial dataset
dataset.tmp <- as.tbl(Stations_covariates_wgs84@data)
dataset.tmp
dim(dataset.tmp)
summary(dataset.tmp)
names(dataset.tmp)

# Clean dataset
names(dataset.tmp) <- gsub("_pt", "", names(dataset.tmp))

# Add a new covariate: bathymetry as class factor covariate
# c("1depth < 5 m", "2depth 5 - 10 m", "3depth 10 - 20 m", "4depth 20 - 50 m")
dataset.orig <- dataset.tmp %>%
  mutate(Bathy_c = 
           cut(-Bathy, breaks = c(min(-Bathy), 5, 10, 20, max(-Bathy, 50)),
               labels = c("1depth < 5m", "2depth 5-10m",
                          "3depth 10-20m", "4depth 20-50m"),
               include.lowest = TRUE)) %>%
  # dplyr::select(-matches("coords"), -optional, -id) %>%
  dplyr::select(lat:Density, Sedim, Bathy, Bathy_c) %>%
  mutate(Year = as.factor(as.character(Year)))

# Extract dataset with only positive values
dataset <- dataset.orig %>%
  filter(Density > 0)

```


### Exploration

*Promis, j'ai bien exploré !*

## Modèle de données positives

```{r}
poly(dataset$Bathy, 3)
# _Additionnal analysis of residuals of the chosen model ----
model <- glm(log(Density) ~ poly(Bathy, 3) + Sedim,
             family = gaussian, data = dataset)
summary(model)
AIC(model) + 2 * sum(log(dataset$Density))
```

### Correction de Laurent

```{r}
dist_gauss <- rnorm(5000, 0, 1)
hist(dist_gauss)
mean(dist_gauss)

exp_gauss <- exp(dist_gauss)
hist(exp_gauss)
mean(exp_gauss)

exp(mean(dist_gauss))
exp(mean(dist_gauss) + 0.5 * var(dist_gauss))

pred_link <- 3 + dist_gauss
hist(pred_link)
pred <- exp(pred_link)
hist(pred)
mean(pred)
exp(3 + 0.5 * var(dist_gauss))
```

## Modèle de données présence-absence

Exécuter tous les chunk précédents : Ctrl + Alt + P
```{r}
# Extract dataset with only positive values
dataset <- dataset.orig %>%
  mutate(Presence = as.numeric(Density > 0))
```

### Exploration
```{r}

# _Covariates potential effect ----
# Histograms (ggplot2)
g <- ggplot(dataset) +
  geom_histogram(aes(Presence),
                 bins = 30, fill = "cyan", col = "grey20")
plot(g)

g + facet_wrap(~Bathy_c)
g + facet_wrap(~Sedim)
g_hist <- g + facet_grid(Sedim ~ Bathy_c)


# Mean by class of factor
ggplot(dataset, aes(Year, Presence)) + 
  stat_summary(fun.y = mean, geom = "bar", fill = "grey70")

ggplot(dataset, aes(Bathy_c, Presence)) + 
  stat_summary(fun.y = mean, geom = "bar", fill = "grey70") +
  facet_wrap(~Sedim)

```

### Ajustement du modèle
```{r}
Dens01.glm <- glm(Presence ~ Bathy + Sedim, family = binomial, data = dataset)

model <- Dens01.glm
AIC(model)
# _Outputs ----
# Anova
summary(model)
anova(model, test = "Chisq")

# % explained deviance
pc_dev_expl <- anova(model)[-1, 2] * 100 / anova(model)[1, 4]
# % explained per factor
print(cbind(rownames(anova(model))[-1], round(pc_dev_expl , digit = 1)))
# % explained per df
print(cbind(rownames(anova(model))[-1], 
            round(pc_dev_expl / anova(model)[-1, 1], digit = 1)))

# Residuals
par(mfrow = c(2, 2))
plot(model, which = c(1, 2, 3, 4))
```

### Sélection sur AUC

```{r}
# _Area under the curve ----
calc.roc <- roc(dataset$Presence, fitted(model))
plot(calc.roc, auc.polygon = TRUE)
auc(calc.roc)

# _Validation of the model on Years 2000-2003 ----
calib.dataset <- dataset[-which(
  dataset$Year %in% c("2000", "2001", "2002", "2003")), ]
valid.dataset <- dataset[which(
  dataset$Year %in% c("2000", "2001", "2002", "2003")), ]
Dens01.valid.glm <- glm(Presence ~ Bathy + Sedim,
                        # family = binomial, 
                        family = binomial(link = "cloglog"),
                        data = calib.dataset)
valid.fit <- predict(Dens01.valid.glm,
                     newdata = valid.dataset, type = "response")

# Area under the curve on validation
valid.roc <- roc(valid.dataset$Presence, valid.fit)
plot(valid.roc, auc.polygon = TRUE)
auc(valid.roc)

```


